---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fpcaCor

We want to focus on FPC analysis from a different angle. Now we consider given $X_i(t)$, $t=1, \dots, T$, $i = 1, \dots, n$. For simplicity, we will assume that the data start with the same time points $t$ across subjects and equi-distant time points. Instead of covariance matrix, we want to work on sample correlation matrix $\hat{K} \in \mathbb{R}^{T \times T}$ based on latent Gaussian copulas. Then by finding appropriate method to smooth $\tilde{K}$, we will obtain the eigenvalues. This is the main goal of the project.

## Installation

You can install the development version of fpcaCor from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("gozdesert/fpcaCor")
```

## Example

```{r example}
library(fpcaCor)
## basic example code
```

